---
title: "HW 10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
accident <- read.csv('https://raw.githubusercontent.com/yumouqiu/DS202-Spring2023/main/Practice/data/fars2016/accident.csv')

person <- read.csv('https://raw.githubusercontent.com/yumouqiu/DS202-Spring2023/main/Practice/data/fars2016/person.csv')

vehicle <- read.csv('https://raw.githubusercontent.com/yumouqiu/DS202-Spring2023/main/Practice/data/fars2016/vehicle.csv')
```
1. are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?
```{r}
accident
#counts number of accidents on each day of the week
day_counts <- table(accident$DAY_WEEK)
most <- names(day_counts)[which.max(day_counts)]
most
```
The seventh day of the week has the most accidents

2. what time of the day do accidents happen (use variable HOUR)?
```{r}
time_counts <- table(accident$HOUR)
timeofday <- names(time_counts)[which.max(time_counts)]
timeofday
```
The 18th hour, aka 6 pm

3. what is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?
```{r}
drunkdriver <- sum(accident$DRUNK_DR >= 1)
drunkdriver
```
8,720 accidents

4. Connect to the person table. Identify drivers (PER_TYP == 1, see fars manual ) and subset on them.

5. Join accident and driver table (work out which variable(s) to use)
6. Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results!